---
title: "Gapminder Analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

## Gather data
#download file,need to specify url and destination of file
```{r}
download.file(url="https://raw.githubusercontent.com/swcarpentry/r-novice-gapminder/gh-pages/_episodes_rmd/data/gapminder_data.csv", destfile="data/gapminder-data.csv")
```
```{r}
read.csv("data/gapminder-data.csv")
```

```{r}
download.file(url="https://raw.githubusercontent.com/swcarpentry/r-novice-gapminder/gh-pages/_episodes_rmd/data/gapminder_data.csv", destfile="data/gapminder_data.csv")
gapminder <- read.csv("data/gapminder_data.csv")
```


```{r}
getwd()
```

```{r}
head(gapminder)

```

## Plotting some data
#ggplot; layers of data plot
illustrator, photoshop
photomanupulation programs allow us to 

### some questions; how does life expectancy change with GDP per capita?

```{r}
library(ggplot2)


```
### how does life expectancy change with GDP per capita?

#mapping; static function
```{r}
ggplot(data = gapminder, mapping = aes(x=gdpPERcap, y=lifeExp))

```


```{r}
ggplot(data = gapminder, mapping = aes(x=gdpPercap, y=lifeExp)) + geom_point()

```

#how does life expectancy change per year?

```{r}
ggplot(data = gapminder, mapping = aes(x=year, y=lifeExp)) + geom_line()

```

```{r}
ggplot(data = gapminder, mapping = aes(x=year, y=lifeExp, by=country)) + 
	geom_line()
	
```

```{r}
ggplot(data = gapminder, mapping = aes(x=year, y=lifeExp, by=country, color=continent)) + 
	geom_line()
	
```

```{r}
ggplot(data = gapminder, mapping = aes(x=year, y=lifeExp, by=country, color=continent)) + 
	geom_line() +
	scale_color_manual(values = c("purple", "orange", "grey", "grey", "grey"))
	
```

```{r}
ggplot(data = gapminder, mapping = aes(x=gdpPercap, y=lifeExp)) + 
	geom_point() +
 	scale_x_log10()+
  
 geom_smooth(method="lm", size=1.5)
```
```{r}
ggplot(data = gapminder, mapping = aes(x=gdpPercap, y=lifeExp)) + 
	geom_point(size=1, alpha=.1) +
 	scale_x_log10()+
  
 geom_smooth(method="lm", size=1.5)
```

```{r}
ggplot(data = gapminder, mapping = aes(x=year, y=lifeExp, by=country)) + 
	geom_line(mapping = aes (color = gapPercap)) + 
  geom_point()
	
```

### How does life expectancy change per year in the americas?

```{R}
 gapminder[gapminder$continent == "Americas",]
```
```{R}
 america <- gapminder[gapminder$continent == "Americas",]
 ggplot(data = america, mapping = aes(x = year, y= lifeExp)) +
   geom_line() +
   facet_wrap(~country)


```

```{R}
 america <- gapminder[gapminder$continent == "Americas",]
 ggplot(data = america, mapping = aes(x = year, y= lifeExp)) +
   geom_line() +
   facet_wrap(~country)+
   theme(axis.text.x = element_text(angle=45)) #x axis angle was changed

```

```{r}
america <- gapminder[gapminder$continent == "Americas",]
 ggplot(data = america, mapping = aes(x = year, y= lifeExp, color = country)) +
   geom_line()
```
# how to save

```{r}
 america <- gapminder[gapminder$continent == "Americas",]
    lifeExp_plot <- ggplot(data = america, mapping = aes(x = year, y= lifeExp)) +
    geom_line()+
    facet_wrap(~country)+
   theme(axis.text.x = element_text(angle=45)) 

    lifeExp_plot
    
```

```{r}
ggsave(plot=lifeExp_plot, filename="lifeExp.png")
```


```{r}
ggplot(data = gapminder, mapping = aes(x = year, y= lifeExp, color =  continent)) +
    geom_line(aes(by = country, alpha=0.2))+
    stat_smooth(aes(by = continent), size=2)
```
    